# loading the data into variable

```{r}
library(plyr)
library(ggplot2)
activity<-read.csv("./repdata_data_activity/activity.csv",colClasses = c("integer","Date","integer"))
stepsperday<-ddply(activity,c("date"),summarise,totalsteps=sum(steps,na.rm=TRUE))
stepsper5min<-ddply(activity,c("interval"),summarise,meansteps=mean(steps,na.rm=TRUE))
```


##What is mean total number of steps taken per day ?

## We have two steps here ,Mean total number of steps taken perday is  ##mean(stepsperday$totalsteps,na.rm=TRUE) and median number
## of steps taken is median(stepsperday$totalsteps)..So
```{r}
stepshist<-ggplot(stepsperday,breaks=c(30,50,30,50),aes(x=totalsteps,col='red'))+geom_histogram()+
   xlab("Total number of steps")+ggtitle("Histogram of total steps in one day")+theme_classic()
print (stepshist)
```


##What is the average daily activity pattern ?

```{r}
daytime<-ggplot(stepsper5min,aes(x=interval,y=meansteps,colour='red'))+geom_line(colour='red',size=1)+
ggtitle("Average steps for each 5-min interval")+ylab("Mean steps")+ theme_bw()
print (daytime)
```


## On average for all days in the dataset ,the five minute intervals contains the following maximum number of steps :
stepsper5min[which.max(stepsper5min$meansteps),]

##*****Imputting missing values

```{r}
hist(which(complete.cases(activity)),col='brown',main="Count of complete cases ",
  xlab="observation number", ylab="Count of complete cases" )
```

##Interpolation is done by using average of the previous observation,the average for the relevant 5-min interval 

```{r}
interpolation<-function(rownumber) { prevrow=rownumber;
  nextrow=rownumber;
while(is.na(activity$steps[prevrow])) {
prevrow=prevrow-1

if (prevrow<1)return(mean(activity[activity$interval==activity$interval[rownumber],"steps"],na.rm=TRUE))
}
while(is.na(activity$steps[nextrow])) { nextrow=nextrow+1
     
if(nextrow>nrow(activity))return(mean(activity[activity$interval==activity$interval[rownumber],"steps"],na.rm=TRUE))
 }
   return (
    (activity$steps[prevrow]+activity$steps[nextrow])/2
)
}

activity_guessNA<-activity
for(n in 1:nrow(activity)) {
if(is.na(activity$steps[n])) {
activity_guessNA$steps[n]=interpolation(n);
}
}
```



## I know , this iss a density plot not a histogram, but the meaning is the same and I didn't want to superimpose two histograms.
##Most of tthe imputed entries appear to haave been added in the most commonly occuring range.

```{r}
stepsperday2<-merge( ddply(activity_guessNA,c("date"),summarise,guesstotalsteps=sum(steps,na.rm=TRUE) ),
stepsperday, by= "date"
)

guesscheck<-ggplot(stepsperday2,aes(x=totalsteps))+ geom_density()+geom_density(aes(x=guesstotalsteps,color="Imputed"))+
ggtitle("density plot comparing raw and NA imputed activity datasets")+xlab("total steps")+ theme_bw()
print(guesscheck)

```



##Here's the histogram for fellow pedants:


```{r}
forpeoplewhoreallywanttoseeahistogram<- ggplot(stepsperday2,aes(x=guesstotalsteps))+ geom_histogram()+ggtitle("Histogram of total number of steps per day after missing values imputed")+theme_bw()
print(forpeoplewhoreallywanttoseeahistogram)

```



##***Are there differences in activity pattern between weekdays and weekends ?
```{r}
interpolation<-function(rownumber) { prevrow=rownumber;
  nextrow=rownumber;
while(is.na(activity$steps[prevrow])) {
prevrow=prevrow-1

if (prevrow<1)return(mean(activity[activity$interval==activity$interval[rownumber],"steps"],na.rm=TRUE))
}
while(is.na(activity$steps[nextrow])) { nextrow=nextrow+1
     
if(nextrow>nrow(activity))return(mean(activity[activity$interval==activity$interval[rownumber],"steps"],na.rm=TRUE))
 }
   return (
    (activity$steps[prevrow]+activity$steps[nextrow])/2
)
}

activity_guessNA<-activity
for(n in 1:nrow(activity)) {
if(is.na(activity$steps[n])) {
activity_guessNA$steps[n]=interpolation(n);
}
}



```




##  Here is Density plot not a histogram, but the meaning is the same and I didn't want to superimpose two histograms.
##Most of the imputed entries is appear to have been added in the most commonly occuring .

```{r}
stepsperday2<-merge( ddply(activity_guessNA,c("date"),summarise,guesstotalsteps=sum(steps,na.rm=TRUE) ),
stepsperday, by="date"
)

guesscheck<-ggplot(stepsperday2,aes(x=totalsteps))+ geom_density()+geom_density(aes(x=guesstotalsteps,color="Imputed"))+
ggtitle("density plot comparing raw and NA imputed activity datasets")+xlab("total steps")+ theme_bw()
print(guesscheck)
```




##Here's the histogram for fellow pedants:

```{r}
forpeoplewhoreallywanttoseeahistogram<- ggplot(stepsperday2,aes(x=guesstotalsteps))+ geom_histogram()+ggtitle("Histogram of total number of steps per day after missing values imputed")+theme_bw()
print(forpeoplewhoreallywanttoseeahistogram)
```



##Are the re differences in activity pattern between weekdays and weekends ?we have to look at the dfay of the week for every single measurement in data that we are analysing.

```{r}
paindays=c("monday","Tuesday","Wednesday","Thurday","Friday") 

activity_guessNA$weekday<-as.factor(ifelse(weekdays(activity_guessNA$date)%in%paindays,"weekday","weekend"))

stepsperinterval.weekdaysplit<-ddply(activity_guessNA,c("interval","weekday"),summarise,meansteps=mean(steps,na.rm=TRUE)
)

weekdayplot<-ggplot(stepsperinterval.weekdaysplit,aes(x=interval,y=meansteps))+facet_wrap(~weekday,nrow=2,ncol=1)+geom_line()+theme_gray()+ggtitle("Mean steps over each 5 min interval split by weekday/weekend")+
  ylab("Mean steps")+
  xlab("Interval number")
print(weekdayplot)
```





